Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-12 VMCAI-revised-version/final.tex
up

 A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
%
%
%
Wqo admits many other equivalent formulations. 
%
%
As an example, $(\N^d, \leq)$, the set of vectors of $d$ natural numbers (where $d$ is finite) with component-wise order is a wqo.
%
Quasi-orderings that have no 
infinite subset of mutually incomparable elements (antichains)
 enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2, \ldots, J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup \ldots \cup J_n)$.
%
%
In what follows, a downward-closed set $D$ is represented by its finite set of ideals (or by the minimal elements of its upward-closed complement), and an upward-closed set $U$ is represented by its finite set of minimal elements. \\
Let us recall a result of Jancar \cite{DBLP:journals/ipl/Jancar99} that proved: that a wqo $(X,\leq)$ is an $\omega^2-wqo$ iff $(Ideals(X),\subseteq)$ is a wqo.

@@

Several families of formal models of processes give rise to WSTSs in a natural way, e.g. Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.

>>>>

 A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
%
%
%
Wqo admits many other equivalent formulations. 
%
%
As an example, $(\N^d, \leq)$, the set of vectors of $d$ natural numbers (where $d$ is finite) with component-wise order is a wqo.
%
Quasi-orderings that have no 
infinite subset of mutually incomparable elements (antichains)
 enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2, \ldots, J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup \ldots \cup J_n)$.
%
%
In what follows, a downward-closed set $D$ is represented by its finite set of ideals (or by the minimal elements of its upward-closed complement), and an upward-closed set $U$ is represented by its finite set of minimal elements. \\

@@

Several families of formal models of processes give rise to WSTSs in a natural way with disfferent compatibilities, e.g. Petri nets when inclusion between markings is used as the well-ordering (WSTS with strong compatibility) and lossy channel systems with the subword ordering (WSTS, compatibility is not strong).
%%%%

Let us recall a result of Jancar \cite{DBLP:journals/ipl/Jancar99} that we will use as a simple definition of an $\omega^2$-wqo: a wqo $(X,\leq)$ is an $\omega^2$-wqo iff $(Ideals(X),\subseteq)$ is a wqo. There exists wqo that are not $\omega^2$-wqo (see the use of the Rado ordering in \cite{DBLP:journals/ipl/Jancar99}).

\begin{definition} \cite{DBLP:journals/corr/abs-1208-4549}
A WSTS  $\mathscr{S}=(S, \rightarrow, \leq)$
is an $\omega^2$-WSTS if $(S, \leq)$ is an $\omega^2$-wqo.
\end{definition}
All naturally occurring WSTS are in fact $\omega^2$-WSTS: for example, Petri nets, VASS, reset/transfer VASS, affine VASS with $d$ counters and $Q$ a finite set of control-states are $\omega^2$-WSTS for the usual ordering ($=\times \leq^d$) on $S=Q \times \N^d$. Similarly, lossy channel systems with $d$ channels and $Q$ a finite set of control-states are $\omega^2$-WSTS
for the ordering $=\times \sqsubseteq^d$ (where $\sqsubseteq$ is the subword ordering on $\Sigma^*$) on $S= Q \times (\Sigma^*)^d$.\\
%%%%%%
